---
title: "Carbon footprint"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE)
```


# Carbon footprint and economy

Impacts of Climate Change disproportionately burdening developing countries. Most of the developing countries have a lower carbon footprint, but they are facing the worst impact of global warming. Bangladesh is also not an exception. 


```{r}
#Install all library
library(dplyr) # provides a set of tools for efficiently manipulating data sets in R
library(ggplot2) # data visualization package
library(plotly) # provides online graphing, analytics, and statistics tools 
library("RColorBrewer") #color palettes
library(choroplethrZip) # for regional map of Bangladesh
library(choroplethr)
library(choroplethrAdmin1)
library(choroplethrMaps)
library(ggthemes)

# Carbon footprint and economic status of a country
library(tmap)
data(World, metro, land)
```


# World population distribution

Most of the world population lives near coastal areas.
This makes a majority of population vulnerable to the impact of climate change.

```{r}
tm_shape(land) +
  tm_raster("elevation", palette = terrain.colors(10)) +
  tm_shape(World) +
  tm_borders("white", lwd = .5) +
  tm_text("iso_a3", size = "AREA") +
  tm_shape(metro) +
  tm_symbols(col = "red", size = "pop2020", scale = .5) +
  tm_legend(show = FALSE)

```

# Impact of global warming

Most of the developing countries have a lower carbon footprint,but they are facing the worst impact of global warming.

```{r}
tmap_mode("view")
tm_shape(World) +
  tm_polygons(c("footprint", "economy")) +
  tm_facets(sync = TRUE, ncol = 2)


# Bangladesh is also not an exception. 
Bangladesh = subset(World, name == "Bangladesh")

tmap_mode("view")
tm_shape(Bangladesh) + 
  tm_polygons(c("footprint", "economy")) +
  tm_facets(sync = TRUE, ncol = 2)

```

# Administrative regions of Bangladesh

Bangladesh is divided into seven administrative regions.

```{r}
data("admin1.map")

bangladesh = subset(admin1.map, admin == "bangladesh")

bd_map <- ggplot(bangladesh) + geom_polygon(aes(long, lat, fill= region, group = group), 
                                color = "black") + theme_map() +  coord_fixed(1.1)+
  ggtitle("Regional Map of Bangladesh")


plot(bd_map)

knitr::opts_chunk$set(echo = TRUE)

```


